---
title: "datachallenge3: Coffee Ratings Data"
author: "menglu zhao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
library(tidyverse)
library(dplyr)
library(purrr)
library(knitr)
library(ggplot2)
library(Hmisc)
```
## github link https://github.com/mezlulu/Datachallenge

## Problem 1 Write a function that returns a table with the mean and standard deviation of the columns aroma through moisture of the coffee ratings data for different categories of coffee. Produce the table for (a) species, (b) country of origin, and (c) binned mean altitude in meters.

```{r}
coffee_clean <- coffee_ratings %>%
  filter(!is.na(altitude_mean_meters)) %>% # remove all the observation that dont have the altitude_mean_meters
  mutate(across(where(is.character), ~replace_na(., "Missing")),  # putting a 'Missing' for NAs
  altitude_binned = cut2(altitude_mean_meters, cuts = c(0, 1000, 2000, 3000, 4000)))  # binning the variable mean altitude by 1000 meters using cut2

group_vars <- c("species", "country_of_origin", "altitude_mean_meters") # these are the variables that we are interested in, and we want to generate a table for these variable.

selected_columns <- c("aroma", "flavor", "aftertaste", "acidity", "body", "balance", "uniformity", "clean_cup", "sweetness", "cupper_points", "moisture") # these are the variable that we want the mean and standard deviation of.

function1 <- function(data, group_var) {
  result <- data %>%
    group_by(across(all_of(group_var))) %>% # !! read the variables as column
    summarise(
      across( # this function apply the same action to multiple columns
        .cols = all_of(selected_columns), # specify the columns for summary
        .fns = list(mean = ~mean(., na.rm = TRUE), sd = ~sd(., na.rm = TRUE)) # a list of functions we want to use, in this case we want mean and standard deviation.
      )
    )
  return(result)
}

summaries <- map(c("species", "country_of_origin", "altitude_binned"), # these are the variables that we are interested in, and we want to generate a table for these variable.
                 ~function1(coffee_clean, .x) %>% # using the function we create
                   kable(caption = paste("Summary for", .x)))  # using kable to create tables
summaries
```

## Problem 2 Write a function that makes a scatterplot of any two variables from the columns aroma through moisture and colors the points by a particular grouping. Make the plot for (1) aroma versus flavor colored by species and (2) balance versus body colored by binned mean altitude. (Hint: Remove outliers to make a more compelling plot!)

```{r}
scatterplot <- function(x_var, y_var, color_group) {
  coffee_clean <- coffee_clean %>%
  filter(between(coffee_clean[[x_var]], quantile(coffee_clean[[x_var]], 0.05), quantile(coffee_clean[[x_var]], 0.95)),
           between(coffee_clean[[y_var]], quantile(coffee_clean[[y_var]], 0.05), quantile(coffee_clean[[y_var]], 0.95))) # avoid the outliers
  remove_ <- function(title) {
  return(gsub("_", " ", title))
  }
   ggplot(coffee_clean, aes_string(x = x_var, y = y_var, color = color_group)) + # plots two variables from a data frame and colors by group
    geom_jitter(alpha = 0.6) +  # use geom_jitter to make a more aesthetically plot
    labs(
      title = paste("Scatterplot of", x_var, "versus", y_var),
      x = x_var,
      y = y_var,
      color = remove_(color_group)
    ) 
}
arguments <- data.frame(x_var = c("aroma", "balance"),  y_var = c("flavor", "body"), color_group = c("species", "altitude_binned")) # parameters for the two plots.

pmap(arguments, scatterplot) 
# same as scatterplot(data=coffee_clean, x_var="aroma", y_var="flavor",color_group= "species")
# scatterplot(data=coffee_clean, x_var="balance", y_var="body",color_group= "altitude_binned") but now using a pmap functions to iterate over a list of function arguments
```

