---
title: "Data challenge 1"
author: "menglu zhao"
date: "September 25, 2023"
output:
  html_document:
    toc: yes
    toc_float: yes
    fig_height: 4.5
    fig_width: 4.5
  pdf_document:
    fig_height: 3.5
    fig_width: 3.5
  word_document:
    toc: no
---

## Link to the repository: <https://github.com/mezlulu/Datachallenge>
```{r options, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, error=TRUE, message=TRUE, warning=TRUE)
```

#### Load in the required libraries
```{r}
library(readr)
library(tidyverse)
library(here)
library(dplyr)
library(tidyr)
library(RColorBrewer)
```

#### Read in the data using here() function
```{r}
state_milk_production <- read_csv(here("data/state_milk_production.csv"))
clean_cheese <- read_csv(here("data/clean_cheese.csv"))
fluid_milk_sales <- read_csv(here("data/fluid_milk_sales.csv"))
milk_products_facts <- read_csv(here("data/milk_products_facts.csv"))
milkcow_facts <- read_csv(here("data/milkcow_facts.csv"))
```

#### Disable scientific notation
```{r}
options(scipen = 999) 
```

## Figure 1
### Make a scatter plot showing the milk production in pounds versus the average number of milk cows using milkcow_facts

```{r}
milkcow_facts %>% #using the milkcow_facts data
ggplot(aes(x = avg_milk_cow_number/1e6, y = milk_production_lbs/1e6,color= year)) + #the x-axis will be the number of milk cows (in millions) and the y-axis will be milk production (in millions of pounds)
  geom_point() + 
  theme_classic() + #plot the data 
  ggtitle("milk production vs. milk cow numbers") + #label the axes and title the plot
  xlab("average milk production (in million pounds)") +
  ylab("average number of milk cows (in million)")
```

## Figure 2
### Plot line plots for the milk production versus the year faceted by region, to examine the milk total milk production in each region over time using state_milk_production

```{r}
new_state_milk_production <- 
state_milk_production %>% #using state_milk_production data
group_by(region, year) %>% #takes the existing table and converts it into a grouped table
summarize(totalp = sum(milk_produced)) #total production of milk each year by region 
  
ggplot(new_state_milk_production, aes(x = year, y = totalp/1e6, color = region)) +
  geom_line() +  #line plots
  labs(x = "Year", y = "Total Milk Production (millions of pounds)", title = "Total Milk Production by Region Over Time") + #label the axes and title the plot
  facet_wrap(~ region, ncol = 2, scales = "free") +   #facet the plot by region
  theme(legend.position = "none")  #remove the legend
```

## Figure 3
### line plot showing the amount (in pounds) of each milk products sold over time using fluid_milk_sales.

```{r}
new_fluid_milk_sales <- 
fluid_milk_sales %>% #using fluid_milk_sales data
data.frame(x=log(fluid_milk_sales$year), y=log(fluid_milk_sales$pounds)) #log scale the data
ggplot(new_fluid_milk_sales, aes(x = x, y = y, color = milk_type)) +
  geom_line() +  #add lines
  labs(x = "Time", y = "Pounds of Product", title = "Milk Product Sales Over Time")  #label the axes & add the title for the plots
```

## Figure 4
### Stacked bar plot of the different type of cheeses consumed over time using a long format version of clean_cheese that you create. Write a caption for this plot.

```{r}
clean_cheese_long <- 
  clean_cheese%>%
  select(-c(`Total American Chese`,`Total Italian Cheese`,`Total Natural Cheese`,`Total Processed Cheese Products`)) #remove all the total variable

new_clean_cheese_long <- 
  pivot_longer(clean_cheese_long, cols=-Year, names_to = "CheeseType", values_to = "Pounds") #change the data to long format and store the data

new_clean_cheese_long%>%
ggplot(aes(x = Year, y = Pounds, fill = CheeseType)) + #plot the data, x is the time and y is the pounds of cheese, color the plot by different types of cheese
  geom_bar(stat = "identity") +
  labs(x = "Time", y = "Pounds of Cheese", title = "cheese consumption Over Time") + #label the axes and title the graph
  scale_fill_brewer(palette = "Paired")  # set color by using RColorBrewer
```

## Figure 5
### A bar plot showing the total Milk Production of different region by different milk type over Time

```{r}
plot5 <- 
  inner_join(new_state_milk_production, fluid_milk_sales, by = "year") #join the two data frames with the common variable "year"
 
ggplot(plot5, aes(x = year, y = pounds/1e9, fill = region)) + 
  geom_bar(stat = "identity") +  #creating bar graph
  labs(x = "Year", y = "Total Milk Production (billion of pounds)", title = "Total Milk Production of different region by different milk type over Time") + #label the axes and title the plot
  facet_wrap(~ milk_type, ncol = 2, scales = "free") + #facet the plot by milk type
  scale_fill_brewer(palette = "Paired") #change the color from the default ggplots2 colors
```
